NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all

### 2023-04-18 SI.tex
Update text

\begin{figure}[!h]
    \centering
    \includegraphics[width = \textwidth]{Figures/LateralDiffusionSI.pdf}
    \caption{Lateral diffusion of water as a function of (A) area per lipid, (B) temperature, (C) membrane thickness, and (D) fraction of charged lipids in a membrane.}
    \label{fig:diffusionSI}
\end{figure}

>>>>

\begin{figure}[!h]
    \centering
    \includegraphics[width = \textwidth]{Figures/LateralDiffusionSI.pdf}
    \caption{Lateral diffusion of water as a function of (A) area per lipid, (B) temperature, (C) membrane thickness, and (D) fraction of charged lipids in a membrane. Non-zero permeation and diffusion values from simulations are shown with blue dots. Histogrammed values are shown with black dots. For the mean value in each bin, average weighted with the simulation lengths was used, and error bars show the standard error of the mean. Only bins with more than one microsecond of data in total were used for water permeation. Only simulations with the temperatures between 300-315\,K were used in D.}
    \label{fig:diffusionSI}
\end{figure}

### 2023-04-18 main.tex
Update text

Cellular membranes contain hundreds of different types of lipid molecules that regulate the membrane properties, morphology, and biological functions~\cite{vanmeer08,Lorent:2020a,Slatter:2016a}. Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}. However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at biological conditions. For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood. Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}. Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}. Such systems exhibit intricate emergent behaviour making inaccuracies more difficult to detect, and accumulation of even modest errors may have a dramatic impact on the conclusions drawn.

In contrast to experimental structural biology, where standard protocols to share and quality-evaluate resolved biomolecular structures are firmly established by the Protein Data Bank (PDB)~\cite{montelione13}, equivalent best practices are yet to be defined for MD simulations. The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\href{https://bioexcel-cv19.bsc.es}{bioexcel-cv19.bsc.es}). However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access are still rare. In particular, tools for automatic quality evaluation of membrane simulations, or training sets for machine learning models of membrane-containing systems, are not yet available. Recent advances using machine learning approaches utilizing publicly available databanks, for example to solve protein structures~\cite{jumper21}, emphasize the increasing importance of such resources. 
%The profound, and sometimes unexpected, effect that such open data sharing and quality evaluation can have on the field cannot be understated. At the very least, open access data will reduce the man and computational hours originating from double work, reducing the environmental impact. Beyond that, by providing the training data, the PDB laid the foundation for the machine learning algorithms now closing into solving the protein folding problem and the strides those machine learning tools have facilitated in the biomedical sciences since.

@@

NMRlipids Databank is a community-driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}. It has been designed to improve the Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} of MD simulation data, most importantly the output trajectories. The NMRlipids Databank is constructed using the NMRlipids project protocol, in which all the content is openly accessible throughout the project~\cite{botan15}. Currently, the NMRlipids Databank contains 726 simulation trajectories with the total length of approximately 0.4\,ms. Single-component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}F, are currently most abundant in the NMRlipids Databank, yet mixtures with up to five lipid types are available. The distribution of lipids among the available simulations, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological relative abundance of different lipid types, with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids from different organisms and organelles~\cite{vanmeer08}. Abbreviations and full names of all lipids present in the Databank are listed in Table~\ref{tab:abbreviations}. Force fields used in simulations cover all the essential parameter sets commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and Table~\ref{tab:ForceFields}, including also united atom and polarizable force fields. Therefore, the averages calculated over the Databank can be considered as mean predictions from available lipid models for an average cell membrane.

The overlay structure of the NMRlipids Databank, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications with minimal investment on new infrastructure. Raw simulation data in the {\it Data layer} can be stored in any publicly available location with long term stability---such as Zenodo (\href{https://zenodo.org}{zenodo.org}), figshare (\href{https://figshare.com}{figshare.com}), or NIRD research data archive (\href{https://archive.norstore.no}{archive.norstore.no})---and with permanent links to the data, such as digital object identifiers (DOIs). 
%or uniform resource names (URNs). 
The {\it Databank layer} (\href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}) is the core of the Databank containing all the relevant information about the simulations: links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, quality evaluation of simulations against experimental data, and the computer programs to create the entries and to analyse the five basic properties extracted from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors, membrane thickness, and equilibration times of principal components). Also the values for these five basic properties are stored in the {\it Databank layer}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{Figures/overlay3.pdf}
    \caption{A) Schematic presentation of the overlay structure used in the NMRlipids Databank. A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
    %b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
    B) Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in five or fewer simulations. 
    C) Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
    D) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
    E) Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
    F) Currently available single lipid component and binary mixtures in the NMRlipids Databank. Colorbar shows the number of available simulations with the darkest green indicating three or more.  
    }
    \label{fig:overlay}
\end{figure}

\subsection{NMRlipids Databank-GUI: graphical access to the MD simulation data}\label{section:access}
NMRlipids Databank-GUI, available at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}, provides easy access to the NMRlipids Databank content through a graphical user interface (GUI). Simulations can be searched based on their molecular composition, force field, temperature, membrane properties, and quality; the search results are ranked based on the simulation quality as evaluated against experimental data. Membranes can be visualized, and properties between different simulations and experiments compared. The NMRlipids Databank-GUI enables rapid surveying of what simulation data is available, selection of the best available simulations for specific systems based on ranking lists, and comparisons of basic properties between different types of membranes. Notably, the GUI enables these operations to be performed by scientists with a wide range of backgrounds---including those who do not necessarily have programming expertise or other means to access MD simulation data.

\subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}
The NMRlipids Databank-API, available at \href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}, provides programmatic access to all simulation data in the NMRlipids Databank. This enables wide range of novel data-driven applications---from construction of machine learning models that predict membrane properties, to automatic analysis of virtually any property across all simulations in the Databank. The flowchart in Fig.~\ref{fig:overlay}D illustrates the practical implementation of such an analysis. First, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files. Simulations can then be automatically analyzed with the help of the README.yaml and the mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank. Finally, the analysis results are stored using the same structure as in the {\it Databank layer}. Examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.

While analysis codes and results for basic membrane properties 
%area per lipid, C–H bond order parameters, X-ray scattering form factors and thickness 
are included in the {\it Databank layer}, unlimited further analyses can be implemented by anyone in separate repositories in the {\it Application layer}. When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:overlay}E. Novel analyses that demonstrate the power of NMRlipids Databank in selecting the best simulation models, analysing rare phenomena, and extending MD simulations to new fields are implemented in an {\it Application-layer} repository located at~\href{https://github.com/NMRlipids/DataBankManuscript}{github.com/NMRlipids/DataBankManuscript}. The related codes are listed in Table~\ref{tab:codesApplication}.

%\begin{figure}[t]
%    \centering
%    \includegraphics[width=88mm]{Figures/flowchart2.pdf}
%    \caption{ A) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
%    B) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
%    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank. The points from the best performing simulations in the NMRlipids databank quality evaluation for each lipid (Figs.~\ref{fig:quality} and~\ref{fig:POPC_POPE_dataSI}) are surrounded by black circles. 
%    }
%    \label{fig:POPCPOPEapls}
%\end{figure}

\subsection{Applications of the NMRlipids Databank}
\subsubsection{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
%
%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. 
%
To minimize the detrimental consequences of artificial MD simulation results for their applications, the quality of lipid bilayer MD simulations has to be carefully assessed~\cite{antila22b}. This can be done, for example, against the C--H bond order parameters from NMR spectroscopy~\cite{bacle21,Wur23} and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}. To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments. Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by calculating the probabilities for each C--H bond order parameter to locate within experimental error ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$). Furthermore, the quality against X-ray scattering experiments ($FF_q$) is estimated as the difference in the experimental and simulated locations of the first form factor minimum. These measures can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}). In addition, the ergodicity of conformational sampling of lipids is estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length. Let us demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.

As Fig.~\ref{fig:quality}A illustrates, predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameters used to characterize cellular membranes, diverge between different force fields. To select the most realistic currently available force fields to describe POPC--POPE membranes, we first rank all simulations in the Databank according to the quality of their C--H bond order parameters  (Fig.~\ref{fig:top50simulations}). From these results, we then pick the force fields occurring in Fig.~\ref{fig:quality}A (that is: force fields for which both POPC and POPE simulations exist in the Databank) and rank them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C). Simulations with $\tau_\mathrm{rel}$ clearly above one are discarded. Because the membrane packing correlates with the average order parameter of the \textit{sn}-1 chain (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A. (Figs.~\ref{fig:quality}E--F and~\ref{fig:POPC_POPE_dataSI} also show direct comparisons with the experimental data for the most relevant simulations; for reference, Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation: POPC bilayer with OPLS3e parameters.) Based on these rankings, Lipid17 and Slipids simulations give the most realistic predictions for a POPC membrane, while simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}). For POPE, on the other hand, GROMOS-CKP and Slipids give the most realistic results, while CHARMM36 and Lipid17 predict membranes that are too packed. In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.

@@

%In summary, the automatic quality evaluation of the NMRlipids databank enables the rapid exploration and selection of the best simulation parameters for specific applications without extensive and tedious manual force field evaluation. This possibility will promote more reliable simulation results for wide range of applications, support force field development, and streamline the parametrization of coarse grained force fields against the best atomistic MD simulations. The automatically extracted quantitative quality measures will be particularly useful to guide automatic parametrization procedures.

%Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.

>>>>

Cellular membranes contain hundreds of different types of lipid molecules that regulate the membrane properties, morphology, and biological functions~\cite{vanmeer08,Lorent:2020a,Slatter:2016a}. Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}. However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at physiologicalbiological conditions. For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood. Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}. Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}. Such systems exhibit intricate emergent behaviour making inaccuracies more difficult to detect, and accumulation of even modest errors may have a dramatic impact on the conclusions drawn.

In contrast to experimental structural biology, where standard protocols to share and quality-evaluate resolved biomolecular structures are firmly established by the Protein Data Bank (PDB)~\cite{montelione13}, equivalent best practices are yet to be defined for MD simulations. The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\href{https://bioexcel-cv19.bsc.es}{bioexcel-cv19.bsc.es}). However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access are still rare. In particular, tools for automatic quality evaluation of membrane simulations, or training sets for machine learning models of membrane-containing systems, are not yet available. Recent advances using machine learning approaches utilizing publicly available databanks, for example to solve protein structures~\cite{jumper21}, emphasize the increasing importance of such resources.

@@

NMRlipids Databank is a community-driven catalogue containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}. It has been designed to improve the Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} of MD simulation data, most importantly the output trajectories. The NMRlipids Databank is constructed using the NMRlipids project protocol, in which all the content is openly accessible throughout the project~\cite{botan15}. Currently, the NMRlipids Databank contains 726 simulation trajectories with the total length of approximately 0.4\,ms. Single-component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}F, are currently most abundant in the NMRlipids Databank, yet mixtures with up to five lipid types are available. The distribution of lipids among the available simulations, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological relative abundance of different lipid types, with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids, depending on organism and organelle~\cite{vanmeer08}. Abbreviations and full names of all lipids present in the Databank are listed in Table~\ref{tab:abbreviations}. Force fields used in simulations cover all the essential parameter sets commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and Table~\ref{tab:ForceFields}, including also united atom and polarizable force fields. Therefore, the averages calculated over the Databank can be considered as mean predictions from available lipid models (average over force field parameters) for an average cell membrane (average over lipid compositions).

The overlay structure of the NMRlipids Databank, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications with minimal investment on new infrastructure. Raw simulation data in the {\it Data layer} can be stored in any publicly available location with long term stability
%figshare (\href{https://figshare.com}{figshare.com}), or NIRD research data archive (\href{https://archive.norstore.no}{archive.norstore.no})---
and with permanent links to the data, such as digital object identifiers (DOIs), such as Zenodo (\href{https://zenodo.org}{zenodo.org}). 
%or uniform resource names (URNs). 
The {\it Databank layer} (\href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}) is the core of the Databank containing all the relevant information about the simulations: links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, quality evaluation of simulations against experimental data, and the computer programs to create the entries and to analyse the five basic properties extracted from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors, membrane thickness, and equilibration times of principal components). Also the values for these five basic properties are stored in the {\it Databank layer}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{Figures/overlay3.pdf}
    \caption{A) Schematic presentation of the overlay structure used in the NMRlipids Databank. A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
    B) Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in five or fewer simulations. 
    C) Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
    D) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
    E) Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
    F) Currently available single lipid component and binary mixtures in the NMRlipids Databank. Colorbar shows the number of available simulations with the darkest green indicating three or more.  
    }
    \label{fig:overlay}
\end{figure}

\subsection{NMRlipids Databank-GUI: graphical access to the MD simulation data}
NMRlipids Databank-GUI, available at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}, provides easy access to the NMRlipids Databank content through a graphical user interface (GUI). Simulations can be searched based on their molecular composition, force field, temperature, membrane properties, and quality; the search results are ranked based on the simulation quality as evaluated against experimental data. Membranes can be visualized, and properties between different simulations and experiments compared. The NMRlipids Databank-GUI enables rapid surveying of what simulation data is available, selection of the best available simulations for specific systems based on ranking lists, and comparisons of basic properties between different types of membranes. Notably, the GUI enables these operations to be performed by scientists with a wide range of backgrounds---including those who do not necessarily have programming expertise or other means to access MD simulation data.

\subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}\label{section:access}
The NMRlipids Databank-API, available at \href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}, provides programmatic access to all simulation data in the NMRlipids Databank. This enables wide range of novel data-driven applications---from construction of machine learning models that predict membrane properties, to automatic analysis of virtually any property across all simulations in the Databank. The flowchart in Fig.~\ref{fig:overlay}D illustrates the practical implementation of such an analysis. First, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files. Simulations can then be automatically analyzed with the help of the README.yaml and the mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank. Finally, the analysis results are stored using the same structure as in the {\it Databank layer}. Examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.

While analysis codes and results for basic membrane properties are included in the {\it Databank layer}, unlimited further analyses can be implemented by anyone in separate repositories in the {\it Application layer}. When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:overlay}E. Novel analyses that demonstrate the power of NMRlipids Databank in selecting the best simulation models, analysing rare phenomena, and extending MD simulations to new fields are implemented in an {\it Application-layer} repository located at~\href{https://github.com/NMRlipids/DataBankManuscript}{github.com/NMRlipids/DataBankManuscript}. The related codes are listed in Table~\ref{tab:codesApplication}.

\subsection{Applications of the NMRlipids Databank}
\subsubsection{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
%
%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. 
%
To minimize the detrimental consequences of artificial MD simulation results for their applications, the quality of lipid bilayer MD simulations has to be carefully assessed~\cite{antila22b}. This can be done, for example, against the C--H bond order parameters from NMR spectroscopy~\cite{bacle21,Wur23} and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}. To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments. Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by first calculating the probabilities for each C--H bond order parameter to locate within experimental error and then averaging the possibilities over different lipid segments ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$). Furthermore, the quality against X-ray scattering experiments ($FF_q$) is estimated as the difference in the experimental and simulated locations of the first form factor minimum. These measures can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}). In addition, the ergodicity of conformational sampling of lipids is estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length. Here we demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.

As Fig.~\ref{fig:quality}A illustrates, predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameters used to characterize cellular membranes, diverge between different force fields. To select the most realistic currently available force fields to describe POPC--POPE membranes, we first rank all simulations in the Databank according to the quality of their C--H bond order parameters  (Fig.~\ref{fig:top50simulations}). From these results, we then pick the force fields occurring in Fig.~\ref{fig:quality}A (that is: force fields for which both POPC and POPE simulations exist in the Databank) and rank them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C). Simulations with $\tau_\mathrm{rel}$ clearly above one (larger than 1.3) are discarded. Because the membrane packing correlates with the average order parameter of the \textit{sn}-1 chain (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A. (Figs.~\ref{fig:quality}E--F and~\ref{fig:POPC_POPE_dataSI} also show direct comparisons with the experimental data for the most relevant simulations; for reference, Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation: POPC bilayer with OPLS3e parameters.) Based on these rankings, Lipid17 and Slipids simulations give the most realistic predictions for a POPC membrane, while simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}). For POPE, on the other hand, GROMOS-CKP and Slipids give the most realistic results, while CHARMM36 and Lipid17 predict membranes that are too packed. In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.

@@
